Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Survival data clean up
===============================================================

Import needed libs
```{r}
library(xlsx)
```

Configure file names and paths
```{r}
# read in the first worksheet from the workbook myexcel.xlsx
# first row contains variable names
rawFile <- "/da/survival_analysis/final_ClinicalData_WORK_SURVIVAL_ANALYSIS_cfm.xlsx"
workDir <- "~/Workspace/research/pcbil" 
pathRawFile <- paste(workDir, rawFile, sep="")

completeUniqueFinalSurvDataFileName <- paste(workDir, "/da/survival_analysis/completeUniqueFinalSurvData.csv", sep="")

realPadsAllFilename <- paste(workDir, "/da/output/realPadsAllPcbil.csv", sep="")

```

Import from final survival data `r pathRawFile`
```{r}
finalSurvData <- read.xlsx2(file = pathRawFile, sheetIndex = 1, header = TRUE, colClasses = c(rep("character", 2), "Date", rep("character", 2), "Date", rep("character", 8), "Date", "character"), na.strings = "")

# Integrity check
str(finalSurvData)
```

Clean up: control and fix field types and values
```{r}
# pad 
sum(is.na(finalSurvData$pad))

# person_number
sum(is.na(finalSurvData$person_number))

# date birth
sum(is.na(finalSurvData$date_birth))

# gender
sum(is.na(finalSurvData$gender))

# chemo
table(finalSurvData$chemo, useNA = "always")
levels(finalSurvData$chemo) <- c(levels(finalSurvData$chemo), "TRUE", "FALSE") 
finalSurvData$chemo[finalSurvData$chemo == "y" | finalSurvData$chemo == "y " | finalSurvData$chemo == "Y"]  <- "TRUE" 
finalSurvData$chemo[finalSurvData$chemo == "n" | finalSurvData$chemo == "n "]  <- "FALSE"
finalSurvData$chemo <- as.logical(finalSurvData$chemo)
table(finalSurvData$chemo, useNA = "always")

# date pad
sum(is.na(finalSurvData$date_pad))

# type probe
levels(finalSurvData$type_probe)
table(finalSurvData$type_probe, useNA = "always")
# Fix a missing value (checked)
finalSurvData$type_probe[finalSurvData$pad == "Hepatocellular Cancer|T1084/12"] <- "b" 
# Fix typo
finalSurvData$type_probe[finalSurvData$type_probe == "r "]  <- "r"
finalSurvData$type_probe <- droplevels(finalSurvData$type_probe)
levels(finalSurvData$type_probe)
table(finalSurvData$type_probe, useNA = "always")

# p_r
levels(finalSurvData$p_r)
table(finalSurvData$p_r, useNA = "always")
finalSurvData$p_r[finalSurvData$p_r == ""]  <- NA
finalSurvData$p_r <- droplevels(finalSurvData$p_r)
levels(finalSurvData$p_r)
table(finalSurvData$p_r, useNA = "always")

# p_t
levels(finalSurvData$p_t)
table(finalSurvData$p_t, useNA = "always")
finalSurvData$p_t[finalSurvData$p_t == ""]  <- NA
finalSurvData$p_t <- droplevels(finalSurvData$p_t)
levels(finalSurvData$p_t)
table(finalSurvData$p_t, useNA = "always")

# p_n
levels(finalSurvData$p_n)
table(finalSurvData$p_n, useNA = "always")
finalSurvData$p_n[finalSurvData$p_n == "" | finalSurvData$p_n == "x"]  <- NA
finalSurvData$p_n <- droplevels(finalSurvData$p_n)
levels(finalSurvData$p_n)
table(finalSurvData$p_n, useNA = "always")

# p_l
levels(finalSurvData$p_l)
table(finalSurvData$p_l, useNA = "always")
finalSurvData$p_l[finalSurvData$p_l == "" | finalSurvData$p_l == "x"]  <- NA
finalSurvData$p_l <- droplevels(finalSurvData$p_l)
levels(finalSurvData$p_l)
table(finalSurvData$p_l, useNA = "always")

# p_v
levels(finalSurvData$p_v)
table(finalSurvData$p_v, useNA = "always")
finalSurvData$p_v[finalSurvData$p_v == "" | finalSurvData$p_v == "x"]  <- NA
finalSurvData$p_v <- droplevels(finalSurvData$p_v)
levels(finalSurvData$p_v)
table(finalSurvData$p_v, useNA = "always")

# p_pn
levels(finalSurvData$p_pn)
table(finalSurvData$p_pn, useNA = "always")
finalSurvData$p_pn <- droplevels(finalSurvData$p_pn)
levels(finalSurvData$p_pn)
table(finalSurvData$p_pn, useNA = "always")

# date death
sum(is.na(finalSurvData$date_death))
```
<p> Control case with pT = 'is'

Control for duplicated amoung unique ID fields: pad, personal number
```{r}

# pad
table(duplicated(finalSurvData$pad))

# personal number
#    check for duplicates
dim(finalSurvData)
table(duplicated(finalSurvData$person_number))
#    there are duplicates, we need to split beween unique and non-unique patients
uniquePatsData <- subset(finalSurvData, !duplicated(finalSurvData$person_number))
dim(uniquePatsData)
table(duplicated(uniquePatsData$person_number))

duplicatedPatsData <- subset(finalSurvData, duplicated(finalSurvData$person_number))
dim(duplicatedPatsData)
table(duplicated(duplicatedPatsData$person_number))

# extract the first occurrences for the repeated from uniquePatsData
repeatedPatsData <- subset(uniquePatsData, person_number %in% duplicatedPatsData$person_number)
dim(repeatedPatsData)
table(duplicated(repeatedPatsData$person_number))

# put together rows for all repeated patients in the same dataset
allRepPatsData <- rbind(repeatedPatsData, duplicatedPatsData)
# and order by personal_number
orderedAllRepPatsData <- allRepPatsData[order(allRepPatsData$person_number), ]

# visual examination of dataset with all repeated patients
orderedAllRepPatsData[, c("pad", "person_number", "type_probe", "date_death")]

# by visual assessment: remove multiple instaces of personal number, which are due: multiple diagnostic probes (tumor samples), tumor heterogeneity (where immunoprofile for the various highly distinct tumor components were recorded). Kept resection probes where both biosy and resection.
duplicates2Remove <- c( "Pancreas Ac|T5900/09_1" , "Intrahepatic Cholangiocarcinoma|T17202/11" , "Gallbladder Ac|T17965/10_2"  , "Ampulla Ac|T1634/12_1"	, "Intrahepatic Cholangiocarcinoma|T6343/12_1", "Hepatocellular Cancer|T1805/11", "Pancreas Ac|T21344/10_2", "Ampulla Ac|T1042/10_2", "Ampulla Ac|T12294/11", "Hepatocellular Cancer|T2493/13_2", "Intrahepatic Cholangiocarcinoma|T2320/06_1", "Intrahepatic Cholangiocarcinoma|T792/08_1", "Pancreas Ac|T19266/09", "Hepatocellular Cancer|K19667/11", "Hepatocellular Cancer|T5784/13", "Intrahepatic Cholangiocarcinoma|T6965/08" , "Intrahepatic Cholangiocarcinoma|T696/08_2", "Intrahepatic Cholangiocarcinoma|T12264/12", "Hepatocellular Cancer|T7201/10_2", "Hepatocellular Cancer|T13365/09_2", "Hepatocellular Cancer|T504/11", "Intrahepatic Cholangiocarcinoma|T8451/06_2", "Perihilary Ac|T1236/06", "Perihilary Ac|T19735/11_2", "Intrahepatic Cholangiocarcinoma|T17615/10", "Perihilary Ac|T20407/10", "Intrahepatic Cholangiocarcinoma|T8275/10_2", "Intrahepatic Cholangiocarcinoma|T5709/12_2", "Intrahepatic Cholangiocarcinoma|T3300/12"	
  )

# integrity check
length(duplicates2Remove)
nrow(orderedAllRepPatsData) - length(duplicates2Remove)
nrow(repeatedPatsData)

# remove repeated tumour profiles from finalSurvData
uniqueFinalSurvData <- subset(finalSurvData, !(pad %in% duplicates2Remove) )

# check for duplicates in cleaned survival data set
dim(uniqueFinalSurvData)
table(duplicated(uniqueFinalSurvData$person_number))
```

Calculate age at diagnosis
```{r}
# function to calculate age: https://github.com/jknowles/eeptools/blob/master/R/age_calc.R
age_calc <- function(dob, enddate=Sys.Date(), units='months', precise=TRUE){
  if (!inherits(dob, "Date") | !inherits(enddate, "Date")){
    stop("Both dob and enddate must be Date class objects")
  }
  start <- as.POSIXlt(dob)
  end <- as.POSIXlt(enddate)
  if(precise){
    start_is_leap <- ifelse(start$year %% 400 == 0, TRUE, 
                            ifelse(start$year %% 100 == 0, FALSE,
                                   ifelse(start$year %% 4 == 0, TRUE, FALSE)))
    end_is_leap <- ifelse(end$year %% 400 == 0, TRUE, 
                          ifelse(end$year %% 100 == 0, FALSE,
                                 ifelse(end$year %% 4 == 0, TRUE, FALSE)))
  }
  if(units=='days'){
    result <- difftime(end, start, units='days')
  }else if(units=='months'){
    months <- sapply(mapply(seq, as.POSIXct(start), as.POSIXct(end), 
                            by='months', SIMPLIFY=FALSE), 
                     length) - 1
    # length(seq(start, end, by='month')) - 1
    if(precise){
      month_length_end <- ifelse(end$mon==1, 28,
                                 ifelse(end$mon==1 & end_is_leap, 29,
                                        ifelse(end$mon %in% c(3, 5, 8, 10), 
                                               30, 31)))
      month_length_prior <- ifelse((end$mon-1)==1, 28,
                                   ifelse((end$mon-1)==1 & start_is_leap, 29,
                                          ifelse((end$mon-1) %in% c(3, 5, 8, 
                                                                    10), 
                                                 30, 31)))
      month_frac <- ifelse(end$mday > start$mday,
                           (end$mday-start$mday)/month_length_end,
                           ifelse(end$mday < start$mday, 
                                  (month_length_prior - start$mday) / 
                                    month_length_prior + 
                                    end$mday/month_length_end, 0.0))
      result <- months + month_frac
    }else{
      result <- months
    }
  }else if(units=='years'){
    years <- sapply(mapply(seq, as.POSIXct(start), as.POSIXct(end), 
                           by='years', SIMPLIFY=FALSE), 
                    length) - 1
    if(precise){
      start_length <- ifelse(start_is_leap, 366, 365)
      end_length <- ifelse(end_is_leap, 366, 365)
      start_day <- ifelse(start_is_leap & start$yday >= 60,
                          start$yday - 1,
                          start$yday)
      end_day <- ifelse(end_is_leap & end$yday >=60,
                        end$yday - 1,
                        end$yday)
      year_frac <- ifelse(start_day < end_day,
                          (end_day - start_day)/end_length,
                          ifelse(start_day > end_day, 
                                 (start_length-start_day) / start_length +
                                   end_day / end_length, 0.0))
      result <- years + year_frac
    }else{
      result <- years
    }
  }else{
    stop("Unrecognized units. Please choose years, months, or days.")
  }
  return(result)
}
# Security check: date_birth, date_pad and date_death are of R type Date
class(uniqueFinalSurvData$date_birth)
class(uniqueFinalSurvData$date_pad)
class(uniqueFinalSurvData$date_death)


#Calculate age at diagnosis
ageAtDiagnosis <- round(age_calc(uniqueFinalSurvData$date_birth, uniqueFinalSurvData$date_pad), digits = 2)

# Add column age at surgery to work data set
uniqueFinalSurvData$age_pad <- ageAtDiagnosis
head(uniqueFinalSurvData[, c("pad", "person_number", "date_birth", "date_pad", "age_pad")])

# Set end date of follow-up: 6-sept-2014
dateEndFollowUp <- as.Date("2014-09-06")

# Calculate survival time (months) and set censoring vector
uniqueFinalSurvData$surv_time[is.na(uniqueFinalSurvData$date_death)] <- round(age_calc(uniqueFinalSurvData$date_pad[is.na(uniqueFinalSurvData$date_death)] , dateEndFollowUp , units="months"), digits = 2)  # alive patients, they have NAN date_death
uniqueFinalSurvData$death_observed[is.na(uniqueFinalSurvData$date_death)] <- 0 # alive patients, they have NAN date_death, censored

uniqueFinalSurvData$surv_time[!(is.na(uniqueFinalSurvData$date_death))] <- round(age_calc(uniqueFinalSurvData$date_pad[!(is.na(uniqueFinalSurvData$date_death))] , uniqueFinalSurvData$date_death[!(is.na(uniqueFinalSurvData$date_death))] , units="months"), digits = 2) # death patients, they have registered date_death
uniqueFinalSurvData$death_observed[!(is.na(uniqueFinalSurvData$date_death))] <- 1 # death patients, they have NAN date_death, observed
# Unfortunately, we couldn't get further censoringinformation (loss to follow up)

# Fix info for one patient who died before pad was signed out
summary(uniqueFinalSurvData$surv_time)
uniqueFinalSurvData[uniqueFinalSurvData$surv_time  < 0 , c("surv_time")] <- 0
summary(uniqueFinalSurvData$surv_time)
```

Import diagnostic data (by cluster and anatomical-based diagnosis) from `r realPadsAllFilename` and merge into survival data set
```{r}
realPadsAll <- read.csv(file = realPadsAllFilename)
head(realPadsAll)
colnames(realPadsAll) <- c("pad", "anon_pad", "clin_diag", "cluster") # modify column names for a neat merge
head(realPadsAll)

completeUniqueFinalSurvData <- merge(uniqueFinalSurvData, realPadsAll, by = "pad")
completeUniqueFinalSurvData <- completeUniqueFinalSurvData[, !(colnames(completeUniqueFinalSurvData) %in% c("comments","na.strings"))]
```

Save the cleaned survival data set
```{r}
write.csv(completeUniqueFinalSurvData, completeUniqueFinalSurvDataFileName, row.names=F, na="")
```