Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Descriptive statistics - by clinical diagnoses
===============================================================

```{r}
suppressMessages(require(psych))
suppressMessages(require(pander))
suppressMessages(require(lawstat))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
data_set_file_imputed <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/imputedPcbilD.csv", sep="")
```

Importing imputed dataset `r data_set_file_imputed` ...
```{r}
pcbilDataWork <- read.csv(file = data_set_file_imputed, row.names = 29, colClasses= c(rep("numeric",27), "factor", "character"), na.strings = "?",quote="\"" )
```

Dataset overview
```{r}
head(pcbilDataWork)
dim(pcbilDataWork)
names(pcbilDataWork)
table(pcbilDataWork$clin_diag)
```

Summary statistics - by marker and clinical diagnosis
```{r}
pcbilNames <- names(pcbilDataWork)

describeBy(pcbilDataWork[1:27], pcbilDataWork$clin_diag)
```
Intepretation of skew and kurtosis: skew > 3 and/or kurtosis > 10 for several markers, indicating non-normal distribution. Non-parametric tests shall be used in case of differential expression analys.

Modified robust Brown-Forsythe Levene-type tests based on the absolute deviations from the
  median - by marker and clinical diagnosis
```{r}
markerNames <- c()
pVals <- c()
for(a_markerName in pcbilNames[1:27]) {
  
  # Test for homogeneity of variance
  markerSubset <- pcbilDataWork[, c(a_markerName, "clin_diag")]   
  levResult <- levene.test(markerSubset[, 1], group= markerSubset[, 2])

  markerNames <- c(markerNames, a_markerName)
  pVals <- c(pVals, levResult$p.value)          
  }
summaryDF <- data.frame(Marker = markerNames, p.Val = pVals)
names(summaryDF)[2] <- "p-value"
pandoc.table(summaryDF, style="grid", justify="right")
```
Interpretation of Levene's Test: p-value < 0.05 for most of markers, wich rejects the null hypothesis of equality of variance between groups (clinical diagnoses). Non-parametric tests shall be used in case of differential expression analys.