Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Unsupervised Hierarchical Cluster Analysis - Exploratory
===============================================================

```{r}
suppressMessages(require(gplots))
suppressMessages(require(amap))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
data_set_file_imputed <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/imputedPcbilD.csv", sep="")
```

Importing imputed dataset `r data_set_file_imputed` ...
```{r}
pcbilDataImputed <- read.csv(file = data_set_file_imputed, row.names = 29, colClasses= c(rep("numeric",27), "factor", "character"), na.strings = "?",quote="\"" )
```

Preparing for clustering...
```{r}
imputedMatrix <- t(as.matrix(pcbilDataImputed[, 1:27]))
color.map <- function(clin_diag) { 
  if (clin_diag == "Hepatocellular Cancer") {
  	"green"
	} else if (clin_diag == "Intrahepatic Cholangiocarcinoma") {
		"blue"
	} else if (clin_diag == "Perihilary Ac") {
		"yellow"
	} else if (clin_diag == "Gallbladder Ac") {
		"purple"
	} else if (clin_diag == "Distal Bile Duct Ac") {
		"deeppink"
	} else if (clin_diag == "Pancreas Ac") {
		"red"
	} else if (clin_diag == "Ampulla Ac") {
		"black"
	} else {
		"white"
	} 
}
patientcolors <- unlist(lapply(pcbilDataImputed$clin_diag, color.map))
```

Clustering and heapmap visualizations - adapted from http://www2.warwick.ac.uk/fac/sci/moac/people/students/peter_cock/r/heatmap/

dist() & hclust()
```{r fig.width=15, fig.height=10}
distances <- c("manhattan", "euclidean", "maximum", "canberra", "binary", "minkowski")
linkages <- c("ward", "single", "average", "complete", "mcquitty", "median", "centroid")
cont <- 0
for(aDistance in distances) {
  for(aLinkage in linkages) {
    mainStr <- paste("dist(", aDistance, "):", "hclust(",aLinkage, ")", sep="")
    heatmap.2(imputedMatrix, col=greenred(100), scale="none", ColSideColors=patientcolors,key= cont == 0, symkey=FALSE, density.info="none", trace="none", cexRow=1.0, cexCol= 0.4, distfun=function(x) dist(x,method=aDistance), main= mainStr, hclustfun = function(x) hclust(x, method = aLinkage) )
    cont <- cont + 1
  }
}
```

Dist() & hclust()
```{r fig.width=15, fig.height=10}
distances <- c("euclidean", "maximum", "manhattan", "canberra", "binary", "pearson", "abspearson", "correlation", "abscorrelation", "spearman", "kendall")
linkages <- c("ward", "single", "average", "complete", "mcquitty", "median", "centroid")
cont <- 0
for(aDistance in distances) {
  for(aLinkage in linkages) {
    mainStr <- paste("Dist(", aDistance, "):", "hclust(",aLinkage, ")", sep="")
    heatmap.2(imputedMatrix, col=greenred(100), scale="none", ColSideColors=patientcolors,key= cont == 0, symkey=FALSE, density.info="none", trace="none", cexRow=1.0, cexCol= 0.4, distfun=function(x) Dist(x,method=aDistance), main= mainStr, hclustfun = function(x) hclust(x, method = aLinkage) )
    cont <- cont + 1
  }
}
```
Work proposal - Dist(pearson, correlation):hclust(average)