Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Survival analysis
===============================================================

Import needed libs
```{r}
library(survival)
library(pastecs)
library(rms)
library(AICcmodavg)
```

Configure file names and paths
```{r}
workDir <- "~/Workspace/research/pcbil" 
completeUniqueFinalSurvDataFileName <- paste(workDir, "/da/survival_analysis/completeUniqueFinalSurvData.csv", sep="")
```

Import complete survival data
```{r}
completeSurvData <- read.csv(file = completeUniqueFinalSurvDataFileName)
completeSurvData$date_birth <- as.Date(completeSurvData$date_birth)
completeSurvData$date_pad <- as.Date(completeSurvData$date_pad)
completeSurvData$date_death <- as.Date(completeSurvData$date_death)

# Remove unclassified-tucs, as this is not a real cluster
table(completeSurvData$cluster, useNA = "always")
completeSurvData[ completeSurvData$cluster == "unclassified-tucs", c("cluster") ] <- NA
completeSurvData$cluster <- droplevels(completeSurvData$cluster)
table(completeSurvData$cluster, useNA = "always")


# Integrity check
str(completeSurvData)
summary(completeSurvData$surv_time) # one patient (109) died before pad
round(stat.desc(completeSurvData$surv_time, basic =  FALSE, norm = TRUE), digits=3)
```

<h1>Significance tests for survival</h1>

<h2>All patients</h2>
```{r}
# Relevel so pancreas is baseline category for cox and hr estimates
completeSurvData$clin_diag <- relevel(completeSurvData$clin_diag, "Pancreas Ac")

#   -All, biopsy vs resection
bx_resPcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ type_probe, data = completeSurvData) 
print(bx_resPcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ type_probe, rho=0, data = completeSurvData)
survdiff(Surv(surv_time, death_observed) ~ type_probe, rho=1, data = completeSurvData)

# Univariate Cox proportional hazards regression model.
probetype_coxModel <- coxph(Surv(surv_time, death_observed) ~ type_probe, data = completeSurvData)
summary(probetype_coxModel)
    
#   -All, by cluster
cluster_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = completeSurvData)
print(cluster_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ cluster, rho=0, data = completeSurvData)
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=1, data = completeSurvData)

# Univariate Cox proportional hazards regression model.
cluster_coxModel <- coxph(Surv(surv_time, death_observed) ~ cluster, data = completeSurvData)
summary(cluster_coxModel)

#   -All, by clin_diag
clin_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = completeSurvData)
print(clin_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = completeSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = completeSurvData)

# Univariate Cox proportional hazards regression model.
clindiag_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = completeSurvData)
summary(clindiag_coxModel)
```

<h2>Biopsy patients</h2>
```{r}
# Obtain and attach data set
pxCompleteSurvData <- subset(completeSurvData, type_probe == "b")

# Relevel so pancreas is baseline category for cox and hr estimates
pxCompleteSurvData$clin_diag <- relevel(pxCompleteSurvData$clin_diag, "Pancreas Ac")

#   -Biopsy patients, by cluster
cluster_px_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = pxCompleteSurvData)
print(cluster_px_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ cluster, rho=0, data = pxCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=1, data = pxCompleteSurvData)

# Univariate Cox proportional hazards regression model.
pxcluster_coxModel <- coxph(Surv(surv_time, death_observed) ~ cluster, data = pxCompleteSurvData)
summary(pxcluster_coxModel)

#   -Biopsy patients, by clin_diag
clin_px_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = pxCompleteSurvData)
print(clin_px_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = pxCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = pxCompleteSurvData)

# Univariate Cox proportional hazards regression model.
pxclindiag_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = pxCompleteSurvData)
summary(pxclindiag_coxModel)
```

<h3>Resection patients</h3>
```{r}
# Obtain and attach data set
resCompleteSurvData <- subset(completeSurvData, type_probe == "r")

# Relevel so pancreas is baseline category for cox and hr estimates
resCompleteSurvData$clin_diag <- relevel(resCompleteSurvData$clin_diag, "Pancreas Ac")

#   -Resected patients, by cluster
cluster_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = resCompleteSurvData)
print(cluster_res_PcbilSurvFit, print.rmean = TRUE)

# log-rank or Mantel-Haenszel test
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=0, data = resCompleteSurvData)
# Peto & Peto modification of the Gehan-Wilcoxon test
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=1, data = resCompleteSurvData)

# Univariate Cox proportional hazards regression model.
clm1 <- coxph(Surv(surv_time, death_observed) ~ cluster, data = resCompleteSurvData)
summary(clm1)
#Test the Proportional Hazards Assumption of a Cox Regression
cox.zph(clm1)
# Assessment of multicollinearity (VIF > 4 indicates problem)
vif(clm1)
# Adjusted Akaike information criterion, Bayesian information criteria. The smaller the AIC or BIC, the better the fit.
AICc(clm1, return.K = FALSE, second.ord = TRUE)
BIC(clm1)

# Multivariate Cox proportional hazards regression model.
clm2 <- coxph(Surv(surv_time, death_observed) ~ cluster + p_t + age_pad, data = resCompleteSurvData)
summary(clm2)
# Test the Proportional Hazards Assumption of a Cox Regression
cox.zph(clm2)
# Assessment of multicollinearity (VIF > 4 indicates problem)
vif(clm2)
# Adjusted Akaike information criterion, Bayesian information criteria. The smaller the AIC or BIC, the better the fit.
AICc(clm2, return.K = FALSE, second.ord = TRUE)
BIC(clm2)

# ANOVA for model comparison
anova(clm1, clm2)

###################################################3
#   -Resected patients, by clin_diag
clin_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = resCompleteSurvData)
print(clin_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = resCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = resCompleteSurvData)

# Univariate Cox proportional hazards regression model.
dgm1 <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = resCompleteSurvData)
summary(dgm1)
#Test the Proportional Hazards Assumption of a Cox Regression
cox.zph(dgm1)
# Assessment of multicollinearity (VIF > 4 indicates problem)
vif(dgm1)
# Adjusted Akaike information criterion, Bayesian information criteria. The smaller the AIC or BIC, the better the fit.
AICc(dgm1, return.K = FALSE, second.ord = TRUE)
BIC(dgm1)

# Multivariate Cox proportional hazards regression model.
dgm2 <- coxph(Surv(surv_time, death_observed) ~ clin_diag + p_t + age_pad, data = resCompleteSurvData)
summary(dgm2)
# Test the Proportional Hazards Assumption of a Cox Regression
cox.zph(dgm2)
# Assessment of multicollinearity (VIF > 4 indicates problem)
vif(dgm2)
# Adjusted Akaike information criterion, Bayesian information criteria. The smaller the AIC or BIC, the better the fit.
AICc(dgm2, return.K = FALSE, second.ord = TRUE)
BIC(dgm2)

# ANOVA for model comparison
anova(dgm1, dgm2)

# AIC/BIC are smaller for cluster cox models than for anatomic loation (for the same data), which indicates a better fit (Is it a valid conclussion???)

```


<h3>Resected ICC patients</h3>
```{r}
# Obtain and attach data set
iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Intrahepatic Cholangiocarcinoma")

#   -Resected ICC patients, by cluster
cluster_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = iccResCompleteSurvData)
print(cluster_icc_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ cluster, rho=0, data = iccResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=1, data = iccResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
iccres_coxModel <- coxph(Surv(surv_time, death_observed) ~ cluster, data = iccResCompleteSurvData)
summary(iccres_coxModel)
```


<h3>Resected patients in extrahepatic-pcbil-tuc</h3>
```{r}
# Obtain and attach data set
ehpcbiltucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & cluster == "extrahepatic-pcbil-tuc")

# Relevel so pancreas is baseline category for cox and hr estimates
ehpcbiltucResCompleteSurvData$clin_diag <- relevel(ehpcbiltucResCompleteSurvData$clin_diag, "Pancreas Ac")

#   -Resected -tuc patients, by clin_diag
clin_ehpcbiltuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = ehpcbiltucResCompleteSurvData)
print(clin_ehpcbiltuc_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = ehpcbiltucResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = ehpcbiltucResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
ehpcbilres_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = ehpcbiltucResCompleteSurvData)
summary(ehpcbilres_coxModel)
```

<h3>Resected patients extrahepatic-pcbil-tuc vs intestinal-tuc</h3>
```{r}
# Obtain and attach data set
int_ehpcbiltucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (cluster == "extrahepatic-pcbil-tuc" | cluster == "intestinal-tuc"))

#   -Resected -tuc patients, by clin_diag
int_ehpcbiltuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = int_ehpcbiltucResCompleteSurvData)
print(int_ehpcbiltuc_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ cluster, rho=0, data = int_ehpcbiltucResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ cluster, rho=1, data = int_ehpcbiltucResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
pcbilint_coxModel <- coxph(Surv(surv_time, death_observed) ~ cluster, data = int_ehpcbiltucResCompleteSurvData)
summary(pcbilint_coxModel)
```

<h3>Resected patients intrahepatic vs perihilar</h3>
```{r}
# Obtain and attach data set
ph_iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Intrahepatic Cholangiocarcinoma" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
ph_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = ph_iccResCompleteSurvData)
print(ph_icc_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = ph_iccResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = ph_iccResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
phicc_coxModel <- coxph(Surv(surv_time, death_observed) ~ cluster, data = ph_iccResCompleteSurvData)
summary(phicc_coxModel)
```

<h3>Resected patients pancreas vs perihilar</h3>
```{r}
# Obtain and attach data set
pdac_phResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Pancreas Ac" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
pdac_ph_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = pdac_phResCompleteSurvData)
print(pdac_ph_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = pdac_phResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = pdac_phResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
pdacphres_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = pdac_phResCompleteSurvData)
summary(pdacphres_coxModel)
```

<h3>Resected patients intrahepatic vs hcc</h3>
```{r}
# Obtain and attach data set
hcc_iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Intrahepatic Cholangiocarcinoma" | clin_diag == "Hepatocellular Cancer"))

#   -Resected ICC patients, by cluster
hcc_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = hcc_iccResCompleteSurvData)
print(hcc_icc_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = hcc_iccResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = hcc_iccResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
hcciccres_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = hcc_iccResCompleteSurvData)
summary(hcciccres_coxModel)
```

<h3>Resected patients gb vs perihilar</h3>
```{r}
# Obtain and attach data set
gb_phResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Gallbladder Ac" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
gb_ph_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = gb_phResCompleteSurvData)
print(gb_ph_res_PcbilSurvFit, print.rmean = TRUE)

survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=0, data = gb_phResCompleteSurvData)
survdiff(Surv(surv_time, death_observed) ~ clin_diag, rho=1, data = gb_phResCompleteSurvData)

# Univariate Cox proportional hazards regression model.
gbphres_coxModel <- coxph(Surv(surv_time, death_observed) ~ clin_diag, data = gb_phResCompleteSurvData)
summary(gbphres_coxModel)
```