Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Analysis of missing data & imputation
===============================================================

```{r}
suppressMessages(require(Amelia))
suppressMessages(require(impute))
```

Dataset raw file
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
data_set_file_raw <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/raw/hcc_pcbil_raw.csv", sep="")
imputedPcbilFileName  <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/imputedPcbilD.csv", sep="")
```

Importing dataset `r data_set_file_raw` ...
```{r}

pcbilDataRaw <- read.csv(file = data_set_file_raw, row.names = 39, colClasses= c(rep("numeric",38), rep("character",2)), na.strings = "?",quote="'" )
```

Creating working dataset...
```{r}
pcbilDataWork <- pcbilDataRaw
```

Processed dataset munging
```{r}
names(pcbilDataWork) <- tolower(names(pcbilDataWork))
names(pcbilDataWork)[39] <- "clin_diagn"
pcbilDataWork$clin_diagn <- as.factor(pcbilDataWork$clin_diagn)
```

Let's see how it look like now...
```{r}
head(pcbilDataWork[1:5])
```

Report basic dataset information
```{r}
dim(pcbilDataWork)
names(pcbilDataWork)
```

Composition of the dataset's Y
```{r}
table(pcbilDataWork$clin_diagn)
```


Summary of missing values - per marker
```{r}
propmiss <- function(dataframe) {
  m <- sapply(dataframe, function(x) {
		data.frame(
			nmiss=sum(is.na(x)), 
			n=length(x), 
			propmiss=round(sum(is.na(x))/length(x),2)
		)
	})
	d <- data.frame(t(m))
	d <- sapply(d, unlist)
	d <- as.data.frame(d)
	d$variable <- row.names(d)
	row.names(d) <- NULL
	d <- cbind(d[ncol(d)],d[-ncol(d)])
	return(d[order(d$propmiss), ])
}

reportmiss <- function(dataframe) {
  propMiss <- propmiss(dataframe)
  print(propMiss)
  totalNumValues <- dim(dataframe)[1] * (dim(dataframe)[2] - 1)
  totalMissingValues <- sum(propMiss$nmiss)
  return ( (100 / totalNumValues) * totalMissingValues  )
}

globalPropMissingValues <- reportmiss(pcbilDataWork)
print(paste("Total % missing data in processed dataset: ", round(globalPropMissingValues, 0) , "%", sep=""))
```

Missing values - per row
```{r}

reportmissperrow <- function(dataframe) {

  pads <- c()
  numMissingValues <- c()
  propMissingValues <- c()
  
  aRow <- 1
  while(aRow <= nrow(dataframe)) {
    pads <- c(pads, row.names(dataframe)[aRow])
    numMissing <- sum(is.na(dataframe[aRow,]))
    propMissingValues <- c(propMissingValues, round((100 / (dim(dataframe)[2] -1)) *numMissing,2))
  	numMissingValues <- c(numMissingValues, numMissing)
  	aRow <- aRow + 1
  }
  missingD <- data.frame(pad = pads, nmissing = numMissingValues, propmissing = propMissingValues, stringsAsFactors=F)
  missingD$propmiss_interv <- cut(missingD$propmissing, c(0,10,20,30,40,50,60,70,80,90,100))
  
  print(table(missingD$propmiss_interv))
  
  return (subset(missingD, propmissing > 50)[, 1])
}

numCasesMissGreat50 <- reportmissperrow(pcbilDataWork)
print(paste("Number of cases in work dataset with more than 50% missing data: ", length(numCasesMissGreat50), sep=""))
```

Amelia Missingness Map
```{r fig.width=11, fig.height=8}
missmap(pcbilDataWork)
```

Work proposal - remove before imputation those markers with > 40% missing values:
ttf1, cd146, cd146_nucl, ngfr_str, calretinin, glypican_3, ezh2, synap, hbme1, mesothelin, hepatocyte

Let's create the marker-filtered dataset
```{r}
pcbilColumnFilteredD <- subset(pcbilDataWork, select=-c(ttf1, cd146, cd146_nucl, ngfr_str, calretinin, glypican_3, ezh2, synap, hbme1, mesothelin, hepatocyte))
```

Check the new status of missing values in the marker-filtered dataset
```{r fig.width=11, fig.height=8}
globalFilteredPropMissingValues <- reportmiss(pcbilColumnFilteredD)
print(paste("Total % missing data in filtered dataset: ", round(globalFilteredPropMissingValues, 0) , "%", sep=""))
missmap(pcbilColumnFilteredD)
```

Filtering of markers with >40% missing data resulted in `r round(globalPropMissingValues - globalFilteredPropMissingValues, 0)`% improvement of global data coberture.

Missing values in marker-filtered dataset - per row
```{r}
fCasesMissGreat50 <- reportmissperrow(pcbilColumnFilteredD)
print(paste("Number of cases in filtered dataset with more than 50% missing data: ", length(fCasesMissGreat50), sep=""))
print(fCasesMissGreat50)
# a <- impute.knn(as.matrix(pcbilColumnFilteredD[1:27]))
```

Work proposal - remove before imputation those rows with > 50% missing values.
```{r}
pcbilRowColumnFilteredD <- pcbilColumnFilteredD[!rownames(pcbilColumnFilteredD) %in% fCasesMissGreat50, ]
```

Check the new status of missing values in the row- and marker-filtered dataset
```{r fig.width=11, fig.height=8}
globalFilteredPropMissingValues <- reportmiss(pcbilRowColumnFilteredD)
print(paste("Total % missing data in row- and column-filtered dataset: ", round(globalFilteredPropMissingValues, 0) , "%", sep=""))
missmap(pcbilRowColumnFilteredD)
```

Missing values in rown- and marker-filtered dataset - per row
```{r}
reportmissperrow(pcbilRowColumnFilteredD)
```

Filtering of markers with >40% and rows with >50% missing data resulted in `r round(globalPropMissingValues - globalFilteredPropMissingValues, 0)`% improvement of global data coberture.

Finally, we're ready for KNN-based imputation
```{r}
imputedPcbil <- impute.knn(as.matrix(pcbilRowColumnFilteredD[1:27]))
imputedPcbilD <- as.data.frame(imputedPcbil$data)
imputedPcbilD$clin_diag <- pcbilRowColumnFilteredD$clin_diagn
```

Check the new status of missing values in the row- and marker-filtered dataset
```{r fig.width=11, fig.height=8}
globalFilteredPropMissingValues <-reportmiss(imputedPcbilD)
print(paste("Total % missing data in imputed dataset: ", round(globalFilteredPropMissingValues, 0) , "%", sep=""))
missmap(imputedPcbilD)
```

Save imputed dataset
```{r}
imputedPcbilD$pad <- rownames(imputedPcbilD)
write.csv(imputedPcbilD, imputedPcbilFileName, row.names=F, na="")
```

End of missing data handling :)