Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Clustering description & differential expression analysis
========================================================


```{r}
suppressMessages(require(pander))
suppressMessages(require(multtest))
suppressMessages(require(RankProd))
suppressMessages(require(samr))
```

File names: clustered dataset
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
clusteredPcbilFileName <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/clusteredPcbilD.csv", sep="")
```

Importing clustered dataset `r clusteredPcbilFileName` ...
```{r}
pcbilDataClustered <- read.csv(file = clusteredPcbilFileName, row.names = 30, colClasses= c(rep("numeric",27), rep("factor",2), "character"), na.strings = "?",quote="\"" )
```

```{r}
alpha <- 0.001

tblFun <- function(x, colName){
    tbl <- table(x)
    res <- cbind(tbl,round(prop.table(tbl)*100,2))
    res.df <- cbind(rownames(res), as.data.frame(res))
    rownames(res.df) <- NULL
    colnames(res.df) <- c(colName, 'Count','Percentage')
    res.df
}
```

Clinical diagnoses - cluster membership
```{r}
for(clDg in unique(pcbilDataClustered$clin_diag)) {
  print(clDg)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$clin_diag == clDg, 29], 'cluster')
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Cluster composition
```{r}
for(clName in unique(pcbilDataClustered$cluster)) {
  print(clName)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$cluster == clName, 28], 'clin_diag')  
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Differential expresion analysis - multtest
```{r fig.width=11, fig.height=8}
#diffDiagn1 <- c("intrahepatic-ccc-tuc")
#diffDiagn2 <- c("hcc-tuc")

diffDiagn1 <- c("extrahepatic-pcbil-tuc", "extrahepatic-pcbil-tuc", "intrahepatic-ccc-tuc")
diffDiagn2 <- c("intrahepatic-ccc-tuc", "intestinal-tuc", "hcc-tuc")
seeds  <- c(1097695, 434343,12345)
multtestResults <- list()

cont <- 1
while(cont <= length(diffDiagn1)){

  #if(cont == 3) {
  #  bootn = 1000
  #} else {
  #  bootn = 1000
  #}
  
  pcbilTucs <- subset(pcbilDataClustered, pcbilDataClustered$cluster == diffDiagn1[cont] | pcbilDataClustered$cluster == diffDiagn2[cont])
  
  m <- MTP(X = t(pcbilTucs[1:27]), Y = pcbilTucs$cluster, test="t.twosamp.unequalvar", robust=TRUE, seed = seeds[cont]) # , B = bootn
  
  plot(m)
  
  mresult <- data.frame(m@rawp[m@adjp <= alpha], m@adjp[m@adjp <= alpha])
  mresult <- cbind(rownames(mresult), mresult)
  meansDiffDiagn1 <- c()
  meansDiffDiagn2 <- c()
  meanDiffs <- c()
  for(marker in mresult[,1]) {
    meansDiffDiagn1 <- c(meansDiffDiagn1, mean(pcbilTucs[pcbilTucs$cluster==diffDiagn1[cont], marker]))
    meansDiffDiagn2 <- c(meansDiffDiagn2, mean(pcbilTucs[pcbilTucs$cluster==diffDiagn2[cont], marker]))
  }
  meanDiffs <- round((meansDiffDiagn1 - meansDiffDiagn2),2)
  mresult <- cbind(mresult, meansDiffDiagn1, meansDiffDiagn2, meanDiffs)
  colnames(mresult) <- c("marker", "raw-pval","adj-pval", "mean class 1", "mean class 2", "mean diffs")
  
  mresult <- mresult[order(-mresult[6]), ]
  rownames(mresult) <- NULL
  
  caption_str <- paste("class 1: ", diffDiagn1[cont], " | class 2: ", diffDiagn2[cont], sep="")
  pandoc.table(mresult, caption= caption_str, style="grid",split.tables=100, justify="right")

  multtestResults <- append(multtestResults, list(as.vector(mresult[,1])))
  
  cont <- cont + 1
}
##
# limma 
# my.df=as.data.frame(imputedMatrix)
#M=my.df
# MA <- new("MAList",list(M=M))

```

Differential expresion analysis - rankprod
```{r fig.width=11, fig.height=8}
diffDiagn1 <- c("extrahepatic-pcbil-tuc", "extrahepatic-pcbil-tuc", "intrahepatic-ccc-tuc")
diffDiagn2 <- c("intrahepatic-ccc-tuc", "intestinal-tuc", "hcc-tuc")
seeds  <- c(1097695, 434343,12345)
rankProdResults <- list()

cont <- 1
while(cont <= length(diffDiagn1)){

  pcbilTucs <- subset(pcbilDataClustered, pcbilDataClustered$cluster == diffDiagn1[cont] | pcbilDataClustered$cluster == diffDiagn2[cont])
    
  subsetClasses <- pcbilTucs$cluster == diffDiagn1[cont]  
  subsetClasses <- as.numeric(subsetClasses)
  
  RP.out <- RP(t(pcbilTucs[1:27]), subsetClasses, num.perm=100, logged=FALSE, plot=FALSE, rand=seeds[cont])
  
  plotRP(RP.out, cutoff= alpha)

  tg <- topGene(RP.out,cutoff= alpha,method="pfp",logged=FALSE, gene.names=colnames(pcbilTucs)[1:27])
  print(tg)
  print(paste("class 1: ", diffDiagn1[cont], " | class 2: ", diffDiagn2[cont], sep=""))
  
  rankProdResults <- append(rankProdResults, list(c(rownames(tg[[1]]), rownames(tg[[2]]))))
  
  cont <- cont + 1
}
```


Differential expresion analysis - sam
```{r fig.width=11, fig.height=8}
diffDiagn1 <- c("extrahepatic-pcbil-tuc", "extrahepatic-pcbil-tuc", "intrahepatic-ccc-tuc")
diffDiagn2 <- c("intrahepatic-ccc-tuc", "intestinal-tuc", "hcc-tuc")
seeds  <- c(1097695, 434343,12345)
samResults <- list()

cont <- 1
while(cont <= length(diffDiagn1)){

  pcbilTucs <- subset(pcbilDataClustered, pcbilDataClustered$cluster == diffDiagn1[cont] | pcbilDataClustered$cluster == diffDiagn2[cont])
    
  subsetClasses <- pcbilTucs$cluster == diffDiagn1[cont]  
  subsetClasses <- as.numeric(subsetClasses)
  subsetClasses <- subsetClasses + 1
  
  samfit <- SAM(t(pcbilTucs[1:27]), subsetClasses, resp.type="Two class unpaired", testStatistic = "wilcoxon", regression.method= "ranks" ,nperms = 1000, genenames = colnames(pcbilTucs)[1:27], fdr.output= alpha)
  
  plot(samfit)
  print(samfit)  
  print(paste("class 1: ", diffDiagn1[cont], " | class 2: ", diffDiagn2[cont], sep=""))
  
  samResults <- append(samResults, list(c(samfit$siggenes.table$genes.up[,1], samfit$siggenes.table$genes.lo[,1])) )
  
  cont <- cont + 1
}
```

Differential expresion analysis - common significant markers (intersection) in multtest + rankprod + sam 
```{r}
diffDiagn1 <- c("extrahepatic-pcbil-tuc", "extrahepatic-pcbil-tuc", "intrahepatic-ccc-tuc")
diffDiagn2 <- c("intrahepatic-ccc-tuc", "intestinal-tuc", "hcc-tuc")

cont <- 1
while(cont <= length(diffDiagn1)){
  
  temp <- intersect(multtestResults[[cont]], rankProdResults[[cont]])
  final <- intersect(temp, samResults[[cont]])
  
  print(paste(diffDiagn1[cont], " vs ", diffDiagn2[cont], ":", sep=""))
  print(final)
  print("-------------------------------------------------------")
  print("")
    
  cont <- cont + 1
}
```