Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Survival analysis
===============================================================

Import needed libs
```{r}
library(survival)
library(pastecs)
library(rms)
library(ggplot2)
```

Configure file names and paths
```{r}
workDir <- "~/Workspace/research/pcbil" 
completeUniqueFinalSurvDataFileName <- paste(workDir, "/da/survival_analysis/completeUniqueFinalSurvData.csv", sep="")
```

Import complete survival data
```{r}
completeSurvData <- read.csv(file = completeUniqueFinalSurvDataFileName)
completeSurvData$date_birth <- as.Date(completeSurvData$date_birth)
completeSurvData$date_pad <- as.Date(completeSurvData$date_pad)
completeSurvData$date_death <- as.Date(completeSurvData$date_death)

# Remove unclassified-tucs, as this is not a real cluster
table(completeSurvData$cluster, useNA = "always")
completeSurvData[ completeSurvData$cluster == "unclassified-tucs", c("cluster") ] <- NA
table(completeSurvData$cluster, useNA = "always")


# Integrity check
str(completeSurvData)
summary(completeSurvData$surv_time) # one patient (109) died before pad
round(stat.desc(completeSurvData$surv_time, basic =  FALSE, norm = TRUE), digits=3)
```

<h1>Kaplan-Meier</h1>

<h2>All patients</h2>
```{r}
# KM-curves
#   -All patients
all_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ 1, data = completeSurvData) 
print(all_PcbilSurvFit, print.rmean = TRUE)
plot(all_PcbilSurvFit,  main = "KM - All", xlab = "Time (months)", ylab = "Survival")

#   -All, biopsy vs resection
bx_resPcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ type_probe, data = completeSurvData) 
print(bx_resPcbilSurvFit, print.rmean = TRUE)

survplot(bx_resPcbilSurvFit, xlim = c(0, 150), main = "KM - Biopsy vs resection", xlab = "Time (months)", ylab = "Survival", time.inc = 10, n.risk   = TRUE,  y.n.risk = -0.3, cex.n.risk = 0.6) # Add to shor 6-yrs, xlim = c(0, 72) # Remove CI c("none","bands","bars")[1]

# TODO fix one missing data in probe type


#   -All, by cluster
cluster_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = completeSurvData) # 
print(cluster_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)

#   -All, by clin_diag
clin_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = completeSurvData) 
print(clin_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h2>Biopsy patients</h2>
```{r}
# Obtain and attach data set
pxCompleteSurvData <- subset(completeSurvData, type_probe == "b")

# Biopsy patients
px_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ 1, data = pxCompleteSurvData) 
print(px_PcbilSurvFit, print.rmean = TRUE)

plot(px_PcbilSurvFit,  main = "KM - Biopsy", xlab = "Time (months)", ylab = "Survival")

#   -Biopsy patients, by cluster
cluster_px_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = pxCompleteSurvData)
print(cluster_px_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_px_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)

#   -Biopsy patients, by clin_diag
clin_px_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = pxCompleteSurvData)
print(clin_px_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_px_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resection patients</h3>
```{r}
# Obtain and attach data set
resCompleteSurvData <- subset(completeSurvData, type_probe == "r")

# Resected patients
res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ 1, data = resCompleteSurvData) 
print(res_PcbilSurvFit, print.rmean = TRUE)

plot(res_PcbilSurvFit,  main = "KM - Restion", xlab = "Time (months)", ylab = "Survival")

#   -Resected patients, by cluster
cluster_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = resCompleteSurvData)
print(cluster_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)

#   -Resected patients, by clin_diag
clin_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = resCompleteSurvData)
print(clin_res_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)

```

<h3>Resected HCC patients</h3>
```{r}
# Obtain and attach data set
hccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Hepatocellular Cancer")

#   -Resected HCC patients, by cluster
cluster_hcc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = hccResCompleteSurvData)
print(cluster_hcc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_hcc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected ICC patients</h3>
```{r}
# Obtain and attach data set
iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Intrahepatic Cholangiocarcinoma")

#   -Resected ICC patients, by cluster
cluster_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = iccResCompleteSurvData)
print(cluster_icc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_icc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected Perihilar CCC patients</h3>
```{r}
# Obtain and attach data set
phccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Perihilary Ac")

#   -Resected Perihilar CCC patients, by cluster
cluster_phcc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = phccResCompleteSurvData)
print(cluster_phcc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_phcc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```


<h3>Resected GB patients</h3>
```{r}
# Obtain and attach data set
gbResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Gallbladder Ac")

#   -Resected GB patients, by cluster
cluster_gb_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = gbResCompleteSurvData)
print(cluster_gb_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_gb_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected DPAC patients</h3>
```{r}
# Obtain and attach data set
dpacResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Pancreas Ac")

#   -Resected DPAC patients, by cluster
cluster_dpac_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = dpacResCompleteSurvData)
print(cluster_dpac_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_dpac_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected Ampullary Carcinoma patients</h3>
```{r}
# Obtain and attach data set
ampResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & clin_diag == "Ampulla Ac")

#   -Resected Ampullary Carcinoma patients, by cluster
cluster_amp_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = ampResCompleteSurvData)
print(cluster_amp_res_PcbilSurvFit, print.rmean = TRUE)

survplot(cluster_amp_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients in hcc-tuc</h3>
```{r}
# Obtain and attach data set
hcctucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & cluster == "hcc-tuc")

#   -Resected -tuc patients, by clin_diag
clin_hcctuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = hcctucResCompleteSurvData)
print(clin_hcctuc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_hcctuc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients in icc-tuc</h3>
```{r}
# Obtain and attach data set
icctucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & cluster == "intrahepatic-ccc-tuc")

#   -Resected -tuc patients, by clin_diag
clin_icctuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = icctucResCompleteSurvData)
print(clin_icctuc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_icctuc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients in extrahepatic-pcbil-tuc</h3>
```{r}
# Obtain and attach data set
ehpcbiltucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & cluster == "extrahepatic-pcbil-tuc")

#   -Resected -tuc patients, by clin_diag
clin_ehpcbiltuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = ehpcbiltucResCompleteSurvData)
print(clin_ehpcbiltuc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_ehpcbiltuc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients in intestinal-tuc</h3>
```{r}
# Obtain and attach data set
inttucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & cluster == "intestinal-tuc")

#   -Resected -tuc patients, by clin_diag
clin_inttuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = inttucResCompleteSurvData)
print(clin_inttuc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(clin_inttuc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```


<h3>Resected patients extrahepatic-pcbil-tuc vs intestinal-tuc</h3>
```{r}
# Obtain and attach data set
int_ehpcbiltucResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (cluster == "extrahepatic-pcbil-tuc" | cluster == "intestinal-tuc"))

#   -Resected -tuc patients, by clin_diag
int_ehpcbiltuc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ cluster, data = int_ehpcbiltucResCompleteSurvData)
print(int_ehpcbiltuc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(int_ehpcbiltuc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients intrahepatic vs perihilar</h3>
```{r}
# Obtain and attach data set
ph_iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Intrahepatic Cholangiocarcinoma" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
ph_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = ph_iccResCompleteSurvData)
print(ph_icc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(ph_icc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients pancreas vs perihilar</h3>
```{r}
# Obtain and attach data set
pdac_phResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Pancreas Ac" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
pdac_ph_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = pdac_phResCompleteSurvData)
print(pdac_ph_res_PcbilSurvFit, print.rmean = TRUE)

survplot(pdac_ph_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients intrahepatic vs hcc</h3>
```{r}
# Obtain and attach data set
hcc_iccResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Intrahepatic Cholangiocarcinoma" | clin_diag == "Hepatocellular Cancer"))

#   -Resected ICC patients, by cluster
hcc_icc_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = hcc_iccResCompleteSurvData)
print(hcc_icc_res_PcbilSurvFit, print.rmean = TRUE)

survplot(hcc_icc_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```

<h3>Resected patients gb vs perihilar</h3>
```{r}
# Obtain and attach data set
gb_phResCompleteSurvData <- subset(completeSurvData, type_probe == "r" & (clin_diag == "Gallbladder Ac" | clin_diag == "Perihilary Ac"))

#   -Resected ICC patients, by cluster
gb_ph_res_PcbilSurvFit <- npsurv(formula = Surv(surv_time, death_observed) ~ clin_diag, data = gb_phResCompleteSurvData)
print(gb_ph_res_PcbilSurvFit, print.rmean = TRUE)

survplot(gb_ph_res_PcbilSurvFit, conf = c("none","bands","bars")[1], xlab = "Time (months)", ylab = "Survival", time.inc = 10)
```