Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Descriptive statistics - by cluster
===============================================================

```{r}
suppressMessages(require(psych))
suppressMessages(require(pander))
suppressMessages(require(lawstat))
suppressMessages(require(ggplot2))
suppressMessages(require(gridExtra))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "~/Workspace/research/pcbil/"
clusteredPcbilFileName <- paste(myWorkDirectory, "da/output/clusteredPcbilD.csv", sep="")
```

Importing imputed dataset `r clusteredPcbilFileName` ...
```{r}
pcbilDataClustered <- read.csv(file = clusteredPcbilFileName, row.names = 30, colClasses= c(rep("numeric",27), rep("factor",2), "character"), na.strings = "?",quote="\"" )
```

Function definition
```{r}
tblFun <- function(x, colName){
    tbl <- table(x)
    res <- cbind(tbl,round(prop.table(tbl)*100,2))
    res.df <- cbind(rownames(res), as.data.frame(res))
    rownames(res.df) <- NULL
    colnames(res.df) <- c(colName, 'Count','Percentage')
    res.df
}
```

Dataset overview
```{r}
head(pcbilDataClustered)
dim(pcbilDataClustered)
names(pcbilDataClustered)
```

Cluster sizes
```{r}
table(pcbilDataClustered$cluster)
```

Cluster composition - by clinical diagnoses
```{r}
for(clName in unique(pcbilDataClustered$cluster)) {
  print(clName)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$cluster == clName, 28], 'clin_diag')  
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Clinical diagnoses - cluster membership
```{r}
for(clDg in unique(pcbilDataClustered$clin_diag)) {
  print(clDg)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$clin_diag == clDg, 29], 'cluster')
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Summary statistics - by marker and cluster
```{r}
pcbilNames <- names(pcbilDataClustered)

describeBy(pcbilDataClustered[1:27], pcbilDataClustered$cluster)
```
Intepretation of skew and kurtosis: skew > 3 and/or kurtosis > 10 for many markers, indicating non-normal distribution. Non-parametric tests shall be used in case of differential expression analys.

Histograms
```{r fig.width=12, fig.height=10}
colors <- c("green1", "blue1", "red", "blueviolet")
diagnoses <- c("hcc-tuc", "intrahepatic-ccc-tuc", "extrahepatic-pcbil-tuc", "intestinal-tuc")
cont <- 1
for(a_diagn in diagnoses) {
  tumorSubset <- subset(pcbilDataClustered, cluster == a_diagn)
  plot_list <- list()
  for(marker in pcbilNames[1:27]) {
    pl <- ggplot(tumorSubset, aes_string(x = marker)) + geom_histogram(aes(y = ..density..), fill = colors[cont], colour = "black", binwidth = 10) + labs(x = paste(marker, sep=""), y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(tumorSubset[, marker],na.rm = TRUE), sd = sd(tumorSubset[, marker], na.rm = TRUE)), colour = "blue", size=1) + coord_cartesian(ylim=c(0,1)) + coord_cartesian(xlim=c(0,100)) # + facet_wrap( ~ clin_diag)
    plot_list[[length(plot_list) + 1]] <- pl
  }  
  n <- length(plot_list)
  nCol <- floor(sqrt(n))
  do.call("grid.arrange", c(plot_list, ncol=nCol, main=a_diagn))
  # break
  cont <- cont + 1
}
```

Modified robust Brown-Forsythe Levene-type tests based on the absolute deviations from the
  median - by marker and cluster
```{r}
markerNames <- c()
pVals <- c()
for(a_markerName in pcbilNames[1:27]) {
  
  # Test for homogeneity of variance
  markerSubset <- pcbilDataClustered[, c(a_markerName, "cluster")]   
  levResult <- levene.test(markerSubset[, 1], group= markerSubset[, 2])

  markerNames <- c(markerNames, a_markerName)
  pVals <- c(pVals, levResult$p.value)          
  }
summaryDF <- data.frame(Marker = markerNames, p.Val = pVals)
names(summaryDF)[2] <- "p-value"
pandoc.table(summaryDF, style="grid", justify="right")
```
Interpretation of Levene's Test: p-value < 0.05 for most of markers, wich rejects the null hypothesis of equality of variance between groups (clusters). Non-parametric tests shall be used for differential expression analys.