Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Descriptive statistics - by cluster
===============================================================

```{r}
suppressMessages(require(psych))
suppressMessages(require(pander))
suppressMessages(require(lawstat))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
clusteredPcbilFileName <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/clusteredPcbilD.csv", sep="")
```

Importing imputed dataset `r clusteredPcbilFileName` ...
```{r}
pcbilDataClustered <- read.csv(file = clusteredPcbilFileName, row.names = 30, colClasses= c(rep("numeric",27), rep("factor",2), "character"), na.strings = "?",quote="\"" )
```

Function definition
```{r}
tblFun <- function(x, colName){
    tbl <- table(x)
    res <- cbind(tbl,round(prop.table(tbl)*100,2))
    res.df <- cbind(rownames(res), as.data.frame(res))
    rownames(res.df) <- NULL
    colnames(res.df) <- c(colName, 'Count','Percentage')
    res.df
}
```

Dataset overview
```{r}
head(pcbilDataClustered)
dim(pcbilDataClustered)
names(pcbilDataClustered)
```

Cluster sizes
```{r}
table(pcbilDataClustered$cluster)
```

Cluster composition - by clinical diagnoses
```{r}
for(clName in unique(pcbilDataClustered$cluster)) {
  print(clName)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$cluster == clName, 28], 'clin_diag')  
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Clinical diagnoses - cluster membership
```{r}
for(clDg in unique(pcbilDataClustered$clin_diag)) {
  print(clDg)  
  clSummary <- tblFun(pcbilDataClustered[pcbilDataClustered$clin_diag == clDg, 29], 'cluster')
  pandoc.table(clSummary, style="grid",split.tables=100, justify="right")
}
```                            

Summary statistics - by marker and cluster
```{r}
pcbilNames <- names(pcbilDataClustered)

describeBy(pcbilDataClustered[1:27], pcbilDataClustered$cluster)
```
Intepretation of skew and kurtosis: skew > 3 and/or kurtosis > 10 for many markers, indicating non-normal distribution. Non-parametric tests shall be used in case of differential expression analys.

Modified robust Brown-Forsythe Levene-type tests based on the absolute deviations from the
  median - by marker and cluster
```{r}
markerNames <- c()
pVals <- c()
for(a_markerName in pcbilNames[1:27]) {
  
  # Test for homogeneity of variance
  markerSubset <- pcbilDataClustered[, c(a_markerName, "cluster")]   
  levResult <- levene.test(markerSubset[, 1], group= markerSubset[, 2])

  markerNames <- c(markerNames, a_markerName)
  pVals <- c(pVals, levResult$p.value)          
  }
summaryDF <- data.frame(Marker = markerNames, p.Val = pVals)
names(summaryDF)[2] <- "p-value"
pandoc.table(summaryDF, style="grid", justify="right")
```
Interpretation of Levene's Test: p-value < 0.05 for most of markers, wich rejects the null hypothesis of equality of variance between groups (clusters). Non-parametric tests shall be used for differential expression analys.