Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Unsupervised Hierarchical Cluster Analysis - pValue
===============================================================

```{r}
suppressMessages(require(gplots))
suppressMessages(require(amap))
suppressMessages(require(pvclust))
suppressMessages(require(snowfall))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "/home/bibu/workspace/"
data_set_file_imputed <- paste(myWorkDirectory, "hcc-pcbil_FDATA_ANALYSIS/r/output/imputedPcbilD.csv", sep="")
```

Importing imputed dataset `r data_set_file_imputed` ...
```{r}
pcbilDataImputed <- read.csv(file = data_set_file_imputed, row.names = 29, colClasses= c(rep("numeric",27), "factor", "character"), na.strings = "?",quote="\"" )
```

Preparing for clustering...
```{r}
imputedMatrix <- t(as.matrix(pcbilDataImputed[, 1:27]))
color.map <- function(clin_diag) { 
  if (clin_diag == "Hepatocellular Cancer") {
    "green"
  } else if (clin_diag == "Intrahepatic Cholangiocarcinoma") {
		"blue"
	} else if (clin_diag == "Perihilary Ac") {
		"yellow"
	} else if (clin_diag == "Gallbladder Ac") {
		"purple"
	} else if (clin_diag == "Distal Bile Duct Ac") {
		"deeppink"
	} else if (clin_diag == "Pancreas Ac") {
		"red"
	} else if (clin_diag == "Ampulla Ac") {
		"black"
	} else {
		"white"
	} 
}
patientcolors <- unlist(lapply(pcbilDataImputed$clin_diag, color.map))

###################################
## Function to Color Dendrograms ##
###################################
dendroCol <- function(dend=dend, keys=keys, xPar="edgePar", bgr="red", fgr="blue", pch=20, lwd=1, ...) {
        if(is.leaf(dend)) {
                myattr <- attributes(dend)
                if(length(which(keys==myattr$label))==1){
                  attr(dend, xPar) <- c(myattr$edgePar, list(lab.col=fgr, col=fgr, pch=pch, lwd=lwd))
                	# attr(dend, xPar) <- c(attr$edgePar, list(lab.col=fgr, col=fgr, pch=pch))
                } else {
                	attr(dend, xPar) <- c(myattr$edgePar, list(lab.col=bgr, col=bgr, pch=pch, lwd=lwd))
                }
        }
  return(dend)
}
# Usage: 
# dend_colored <- dendrapply(dend, dendroCol, keys, xPar="edgePar", bgr="red", fgr="blue", pch=20) # use xPar="nodePar" to color tree labels
# plot(dend_colored, horiz=T)
```

Clustering and heapmap visualizations - adapted from http://www2.warwick.ac.uk/fac/sci/moac/people/students/peter_cock/r/heatmap/
Dist() & hclust() - with clininical diagnoses in the side color bar
```{r fig.width=15, fig.height=10}
source("http://faculty.ucr.edu/~tgirke/Documents/R_BioCond/My_R_Scripts/my.colorFct.R") 
hc <- hclust(Dist(t(imputedMatrix), method="pearson"), method="average") # Clusters rows by Pearson
hr <- hclust(Dist(imputedMatrix, method="pearson"), method="average") # Clusters columns by Spearman
heatmap.2(imputedMatrix, Rowv=as.dendrogram(hr), Colv=as.dendrogram(hc), col=greenred(100), scale="none", ColSideColors=patientcolors, density.info="none", trace="none")
```

Dist() & hclust() - with clustering cuts in the side color bar
```{r fig.width=15, fig.height=10}
mycl <- cutree(hc, h=max(hc$height)/2.2)
mycolhc <- sample(rainbow(256))
mycolhc <- mycolhc[as.vector(mycl)]
heatmap.2(imputedMatrix, Rowv=as.dendrogram(hr), Colv=as.dendrogram(hc), col=greenred(100), scale="none", ColSideColors=mycolhc, density.info="none", trace="none")
```

Obtain significant clusters by pvclust bootstrap analysis: adapted from http://manuals.bioinformatics.ucr.edu/home/R_BioCondManual#R_clustering
```{r fig.width=15, fig.height=10}
cl <- makeCluster(2)
pv2 <- parPvclust(cl, imputedMatrix, method.dist="uncentered", method.hclust="average", nboot=25000)
plot(pv2, hang=-1)
pvrect(pv2, alpha=0.95)

clsig <- unlist(pvpick(pv2, alpha=0.90, pv="au", type="geq", max.only=TRUE)$clusters) 
dend_colored <- dendrapply(as.dendrogram(pv2$hclust), dendroCol, keys=clsig, xPar="edgePar", bgr="black", fgr="red", pch=20)

heatmap.2(imputedMatrix, Rowv=as.dendrogram(hr), Colv=dend_colored, col=greenred(100), scale="none", ColSideColors=patientcolors, density.info="none", trace="none") 

heatmap.2(imputedMatrix, Rowv=as.dendrogram(hr), Colv=dend_colored, col=greenred(100), scale="none", ColSideColors=mycolhc, density.info="none", trace="none") 

```