Immunoprofiling of adenocarcinomas of the pancreatobiliary tree
===============================================================
Descriptive statistics - by clinical diagnoses
===============================================================

```{r}
suppressMessages(require(psych))
suppressMessages(require(pander))
suppressMessages(require(lawstat))
suppressMessages(require(ggplot2))
suppressMessages(require(gridExtra))
```

File names: imputed dataset
```{r}
myWorkDirectory <- "~/Workspace/research/pcbil/"
data_set_file_imputed <- paste(myWorkDirectory, "da/output/imputedPcbilD.csv", sep="")
```

Importing imputed dataset `r data_set_file_imputed` ...
```{r}
pcbilDataWork <- read.csv(file = data_set_file_imputed, row.names = 29, colClasses= c(rep("numeric",27), "factor", "character"), na.strings = "?",quote="\"" )
```

Dataset overview
```{r}
head(pcbilDataWork)
dim(pcbilDataWork)
names(pcbilDataWork)
table(pcbilDataWork$clin_diag)
```

Summary statistics - by marker and clinical diagnosis
```{r}
pcbilNames <- names(pcbilDataWork)

describeBy(pcbilDataWork[1:27], pcbilDataWork$clin_diag)
```
Intepretation of skew and kurtosis: skew > 3 and/or kurtosis > 10 for several markers, indicating non-normal distribution. Non-parametric tests shall be used in case of differential expression analys.

Histograms
```{r fig.width=12, fig.height=10}
colors <- c("green", "blue", "yellow", "purple", "deeppink", "red", "black")
diagnoses <- c("Hepatocellular Cancer", "Intrahepatic Cholangiocarcinoma", 
              "Perihilary Ac", "Gallbladder Ac", "Distal Bile Duct Ac", 
              "Pancreas Ac", "Ampulla Ac")
cont <- 1
for(a_diagn in diagnoses) {
  tumorSubset <- subset(pcbilDataWork, clin_diag == a_diagn)
  plot_list <- list()
  for(marker in pcbilNames[1:27]) {
    pl <- ggplot(tumorSubset, aes_string(x = marker)) + geom_histogram(aes(y = ..density..), fill = colors[cont], colour = "black", binwidth = 10) + labs(x = paste(marker, sep=""), y = "Density") + stat_function(fun=dnorm, args=list(mean = mean(tumorSubset[, marker],na.rm = TRUE), sd = sd(tumorSubset[, marker], na.rm = TRUE)), colour = "blue", size=1) + coord_cartesian(ylim=c(0,1)) + coord_cartesian(xlim=c(0,100)) # + facet_wrap( ~ clin_diag)
    plot_list[[length(plot_list) + 1]] <- pl
  }  
  n <- length(plot_list)
  nCol <- floor(sqrt(n))
  do.call("grid.arrange", c(plot_list, ncol=nCol, main=a_diagn))
  # break
  cont <- cont + 1
}
```

Modified robust Brown-Forsythe Levene-type tests based on the absolute deviations from the
  median - by marker and clinical diagnosis
```{r}
markerNames <- c()
pVals <- c()
for(a_markerName in pcbilNames[1:27]) {
  
  # Test for homogeneity of variance
  markerSubset <- pcbilDataWork[, c(a_markerName, "clin_diag")]   
  levResult <- levene.test(markerSubset[, 1], group= markerSubset[, 2])

  markerNames <- c(markerNames, a_markerName)
  pVals <- c(pVals, levResult$p.value)          
  }
summaryDF <- data.frame(Marker = markerNames, p.Val = pVals)
names(summaryDF)[2] <- "p-value"
pandoc.table(summaryDF, style="grid", justify="right")
```
Interpretation of Levene's Test: p-value < 0.05 for most of markers, wich rejects the null hypothesis of equality of variance between groups (clinical diagnoses). Non-parametric tests shall be used in case of differential expression analys.